---
title: "Practical Machine Learning 
author: Kavin

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practical Machine Learning Project

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement ??? a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

# Analysis

Loading data:

```{r load data, warning=FALSE, message=FALSE, echo=TRUE}

training = read.csv("./pml-training.csv",na.strings=c("NA","#DIV/0!",""))
testing = read.csv("./pml-testing.csv",na.strings=c("NA","#DIV/0!",""))

# Data dimensions
dim(training)
dim(testing)
```

```{r first look, warning=FALSE, message=FALSE, eval= FALSE}
# First look at the data
head(training)
head(testing)

```

# Data Collection
Source: [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har)

The training data for this project are available here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The test data are available here:
[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

## Loading the data
Load the files, create 70/30 data partition and check for NAs. Missing data is mapped to NA strings.

```{r}

# Read CSVs (after download, of course) while dealing with NAs
training_data <- read.csv("./pml-training.csv", na.strings = c("NA","","#DIV/0!"))
testing_data <- read.csv("./pml-testing.csv", na.strings = c("NA","","#DIV/0!"))

str(training_data)

```
## Clean-up data
NAs don't help us with predction. So let's go ahead and keep columns without missing values.

```{r}

training_data <- training_data[, colSums(is.na(training_data)) == 0]
testing_data <- testing_data[, colSums(is.na(testing_data)) == 0]

```

## Inspet
We want to reduce noise and irrelevant variables. Let's look at the column names and use our judgement about what to keep and what to ignore.


# Conclusion

The Random Forest algorithm had an accuracy of about 99% whereas the Regression Tree's accuracy was about 70%. The  algorithm outperformed Random Forest by about 30%. 
```
